Queue Aware Precoder Design for Space Frequency Resource Allocation

### 2014-08-21 Report/wsrm_formulation.tex
modified for review
Modifications made for the convergence

The proposed algorithm is referred as \ac{QM} \ac{JSFRA} scheme with a per \ac{BS} power constraint which is outlined in Algorithm \ref{algo-1}. The iterative procedure repeats until the improvement on the objective is less than a predetermined tolerance parameter or the maximum number of iterations is reached. Instead of initializing \me{\tilde{\mbf{u}}_{l,k,n}} arbitrarily to a feasible point, transmit precoders can also be initialized with any feasible point \me{\tilde{\mbf{m}}_{l,k,n}}, which is then used to find \me{\tilde{\mbf{u}}_{l,k,n}} in an efficient manner as briefed in Algorithm \ref{algo-1}. In Algorithm \ref{algo-1}, the \ac{SCA} iterations are carried until convergence or for maximum of \me{I_{\max}} iterations for the optimal \me{\mbf{w}_{l,k,n}} receive beamformers and the outer iterations are for the convergence of the number of queued bits, which is limited by the maximum of \me{J_{\max}} iterations.
\begin{algorithm}
 \SetAlgoLined
 \DontPrintSemicolon
 \BlankLine
 \SetKwInput{KwInit}{Initialize}
 \KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}, \fall n \in \mathcal{C}}}
 \KwOut{\me{\mvec{m}{l,k,n}} and \me{\mvec{w}{l,k,n} \fall l \in \set{1,2,\dotsc,L}}}
 \KwInit{\me{i=0}, \me{j=0} and the transmit precoders \me{\tilde{\mbf{m}}_{l,k,n}} randomly satisfying the total power constraint \eqref{eqn-4.3}}
 update \me{\mvec{w}{l,k,n}} with \eqref{eqn-10} and \me{\tilde{\mbf{u}}_{l,k,n}} with \eqref{eqn-8} using \me{\tilde{\mbf{m}}_{l,k,n}}.\;
 \Repeat{Queue convergence or \me{j \geq J_{\max}}}{
 \Repeat{\ac{SCA} convergence or \me{i \geq I_{\max}}}{
 solve for the transmit precoders \me{\mvec{m}{l,k,n}} using \eqref{eqn-9} \;
 update the constraint set \eqref{eqn-8} with \me{\tilde{p}_{l,k,n}}, \me{\tilde{q}_{l,k,n}} and \me{\tilde{\beta}_{l,k,n}} using \eqref{eqn-wsrm-expr} with the precoders \me{\mvec{m}{l,k,n}} obtained from the previous step. \;
 $i = i + 1$. \;
 }
 update the receive beamformers \me{\mvec{w}{l,k,n}} using \eqref{eqn-9--1} or \eqref{eqn-10} with the recent precoders \me{\mvec{m}{l,k,n}}. \;
 $j = j + 1$. \;
 }
 \caption{Algorithm of \acs{JSFRA} scheme}
 \label{algo-1}
\end{algorithm}

@@

In the proposed solution, we replaced \eqref{eqn-6.2} by a convex constraint using the first order approximation, which basically means that we do not solve the problem exactly. According to the traditional \ac{BCDM}, we need to solve a sub problem when fixing a set of variables to the global optimum to ensure the convergence to a stationary point. If we just approximate the objective, then the convergence is guaranteed \cite{razaviyayn2013unified}. In our case, we solve the sub problem inexactly, so the convergence proof of \ac{BCDM} does not apply to our problem. In this problem, we used alternating optimization with \ac{SCA} method, which provides monotonic convergence since the objective is improved at each step \textit{i.e} \me{f^{(i)} \geq f^{(i+1)}}, assuming \me{f^{(i)}} is the objective function at the \me{i^\mathrm{th}} \ac{SCA} iteration. For a fixed receive beamformers, the problem is guaranteed to converge to a \ac{KKT} point at each \ac{SCA} points as discussed in \cite{marks1978technical}. Now, by alternating the transmit and the receive precoders, we perform \ac{BCDM}, which provides monotonic convergence of the objective. Since the \ac{SCA} method is also considered in our problem, the convergence is not straight forward. We can only speculate the convergence based on the monotonicity of the objective values. Since receive beamformers are based on the \ac{MMSE} objective, it improves the objective value for each update, providing a monotonic increase in the objective value.

>>>>

The proposed algorithm is referred as \acl{QM} \ac{JSFRA} scheme with a per \ac{BS} power constraint which is outlined in Algorithm \ref{algo-1}. The iterative procedure repeats until the improvement on the objective is less than a predetermined tolerance parameter or the maximum number of iterations is reached. Instead of initializing \me{\tilde{\mbf{u}}_{l,k,n}} arbitrarily to a feasible point, transmit precoders can also be initialized with any feasible point \me{\tilde{\mbf{m}}_{l,k,n}}, which is then used to find \me{\tilde{\mbf{u}}_{l,k,n}} in an efficient manner as briefed in Algorithm \ref{algo-1}. In Algorithm \ref{algo-1}, the \ac{SCA} iterations are carried until convergence or for maximum of \me{I_{\max}} iterations for the optimal \me{\mbf{w}_{l,k,n}} receive beamformers and the outer iterations are for the convergence of the number of queued bits, which is limited by the maximum of \me{J_{\max}} iterations.
\begin{algorithm}
 \SetAlgoLined
 \DontPrintSemicolon
 \BlankLine
 \SetKwInput{KwInit}{Initialize}
 \KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}, \fall n \in \mathcal{N}}}
 \KwOut{\me{\mvec{m}{l,k,n}} and \me{\mvec{w}{l,k,n} \fall l \in \set{1,2,\dotsc,L}}}
 \KwInit{\me{i=0}, \me{j=0} and the transmit precoders \me{\tilde{\mbf{m}}_{l,k,n}} randomly satisfying the total power constraint \eqref{eqn-4.3}}
 update \me{\mvec{w}{l,k,n}} with \eqref{eqn-10} and \me{\tilde{\mbf{u}}_{l,k,n}} with \eqref{eqn-8} using \me{\tilde{\mbf{m}}_{l,k,n}}.\;
 \Repeat{Queue convergence or \me{j \geq J_{\max}}}{
 \Repeat{\ac{SCA} convergence or \me{i \geq I_{\max}}}{
 solve for the transmit precoders \me{\mvec{m}{l,k,n}} using \eqref{eqn-9} \;
 update the constraint set \eqref{eqn-8} with \me{\tilde{p}_{l,k,n}}, \me{\tilde{q}_{l,k,n}} and \me{\tilde{\beta}_{l,k,n}} using \eqref{eqn-wsrm-expr} with the precoders \me{\mvec{m}{l,k,n}} obtained from the previous step. \;
 $i = i + 1$. \;
 }
 update the receive beamformers \me{\mvec{w}{l,k,n}} using \eqref{eqn-9--1} or \eqref{eqn-10} with the recent precoders \me{\mvec{m}{l,k,n}}. \;
 $j = j + 1$. \;
 }
 \caption{Algorithm of \acs{JSFRA} scheme}
 \label{algo-1}
\end{algorithm}

@@

In the proposed solution, we replaced \eqref{eqn-6.2} by a convex constraint using the first order approximation, which basically means that we do not solve the problem exactly. According to the traditional \ac{BCDM}, we need to solve a sub problem when fixing a set of variables to the global optimum to ensure the convergence to a stationary point. If we just approximate the objective, then the convergence is guaranteed \cite{razaviyayn2013unified}. In our case, we solve the sub problem inexactly, so the convergence proof of \ac{BCDM} does not apply to our problem. In this problem, we used alternating optimization with \ac{SCA} method, which provides monotonic convergence since the objective is improved at each step \textit{i.e} \me{f^{(i)} \geq f^{(i+1)}}, assuming \me{f^{(i)}} is the objective function at the \me{i^\mathrm{th}} \ac{SCA} iteration. For a fixed receive beamformers, the problem is guaranteed to converge to a \ac{KKT} point at each \ac{SCA} updates as discussed in \cite{marks1978technical}. Now, by alternating the transmit and the receive precoders, we perform \ac{BCDM}, which provides monotonic convergence of the objective.

The linear approximation majorizes the quadratic-over-linear function in \eqref{eqn-6.2} around a fixed point \me{\tilde{\mbf{u}}^{(i)}_{l,k,n}}. The constraint approximation satisfies the following conditions as in \cite{marks1978technical}
\begin{subeqnarray} \label{sca-req}
	f(\tilde{\mbf{u}}_{l,k,n}) &\leq& \bar{f}(\tilde{\mbf{u}}_{l,k,n},\tilde{\mbf{u}}^{(i)}_{l,k,n}) \\
	f(\tilde{\mbf{u}}^{(i)}_{l,k,n}) &=& \bar{f}(\tilde{\mbf{u}}^{(i)}_{l,k,n},\tilde{\mbf{u}}^{(i)}_{l,k,n}) \\
	\partial f(\tilde{\mbf{u}}^{(i)}_{l,k,n}) &=& \partial \bar{f}(\tilde{\mbf{u}}^{(i)}_{l,k,n},\tilde{\mbf{u}}^{(i)}_{l,k,n})
\end{subeqnarray}
where \me{\bar{f}(x,x^{(i)})} is the approximate function majorizing \me{\bar{f}(x)} around the point \me{x^{(i)}}. At each iteration \me{(i)}, the optimal value \me{x^{\ast}} obtained by solving the approximate problem is a feasible solution to next approximation at \me{(i+1)} iteration by following \eqref{sca-req}. Once the problem converged to an optimal point, the receive beamformers are updated based on the receivers mentioned in \eqref{opt-rx} or \eqref{eqn-10}. Since the receiver minimizes the objective value for the fixed transmit precoders, the proposed \ac{JSFRA} scheme guarantees to converge to an optimal point.
